---
title: "Introduction to Machine Learning - HW 1"
author: "maitmad"
date: "2026-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#ASSIGNMENT 1 FOR INTRODUCTIONS TO MACHINE LEARNING

# Using dataset on Aircraft Wildlife Strikes 1990-2015


```{r}
library(readxl)
aircraft_strikes <- read_excel("~/Downloads/aircraft strikes.xlsx")
```

# descriptive statistics for quantitative variables

```{r}
summary(aircraft_strikes$`Incident Year`)  ## quantitative variable referring to year of incident
summary(aircraft_strikes$`Incident Month`) ## quantitative variable referring to month of incident
summary(aircraft_strikes$`Incident Day`)   ## quantitative variable referring to day of incident
```

# descriptive statistics for categorical variables

```{r}
aircraft_strikes$Operator <- as.factor(aircraft_strikes$Operator)
summary(aircraft_strikes$Operator)
table(aircraft_strikes$Operator)
```

#Transformation of variable Incident Year into smaller categories

```{r}
aircraft_strikes$`Incident Year` <- cut(aircraft_strikes$`Incident Year`,
                                         breaks = c(-Inf, 1999, 2010, Inf),
                                         labels = c("1999 or before", "2000 - 2010", "2011 and after"))
table(aircraft_strikes$`Incident Year`)

head(aircraft_strikes$`Incident Year`)
tail(aircraft_strikes$`Incident Year`)
```


#Bar chart of flight phase the incident took place in
```{r}
table(aircraft_strikes$`Flight Phase`)
barplot(table(aircraft_strikes$`Flight Phase`), main = "Aircraft Strikes by Phase of Flight", xlab = "Flight Phase", ylab = "Frequency", col = "blue", las = 3 )
```
#Scatterplot of Incident Year and Incident Month
```{r}
plot(aircraft_strikes$`Incident Year`, aircraft_strikes$`Incident Month`, main = "[Y Variable] vs [X Variable]", xlab = "[Incident Year]", ylab = "[Incident Month]", pch = 12, col = "green")
```
##End of Assignment


